---
title: "assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code for reading in the dataset and/or processing the data


```{r}
data <- read.csv("activity.csv")
head(data)
```
# Histogram of the total number of steps taken each day

```{r}
library(dplyr)
steps_grouped <- data %>% group_by(date) %>% summarise_at(vars(c("steps")), list(steps = sum))
head(steps_grouped)
```


```{r echo=FALSE}
hist(steps_grouped$steps, main="daily steps", breaks=10)
```

# Mean and median number of steps taken each day

```{r}
print("mean")
print(mean(steps_grouped$steps, na.rm = TRUE))
print("median")
print(median(steps_grouped$steps, na.rm = TRUE))
```

# Time series plot of the average number of steps taken

```{r}
library(ggplot2)
steps_avg <- data %>% group_by(date) %>% summarise_at(vars(c("steps")), list(steps = mean))
steps_avg <- steps_avg[!is.na(steps_avg$steps), ]
plt <- ggplot(steps_avg, aes(date, steps, group = 1)) +
         geom_point() +
         geom_line() + 
         labs(x = "Date", y = "Steps per 5 minute") + scale_x_discrete(guide = guide_axis(angle = 90))
plt
```


# The 5-minute interval that, on average, contains the maximum number of steps

```{r}
data_nona <- data[!is.na(data$steps), ]
steps_avg_int <- data_nona %>% group_by(interval) %>% summarise_at(vars(c("steps")), list(steps = mean))
#steps_avg_int <- steps_avg_int[!is.na(steps_avg_int$steps), ]
steps_avg_int <- steps_avg_int[order(steps_avg_int$steps, decreasing = TRUE), ]
steps_avg_int
```
the answer is the interval 835 which corresponds to 13.55 - 14.00 

# Code to describe and show a strategy for imputing missing data
```{r}
sum(is.na(data$steps))
```
There are 2304 rows with NA steps. We will replace them with the mean number of steps on this interval
```{r}
data_merged <- merge(x=data, y=steps_avg_int, by.x='interval', by.y = 'interval') 
head(data_merged)
```


```{r}
data_merged[is.na(data_merged$steps.x), "steps.x"] <- data_merged[is.na(data_merged$steps.x), "steps.y"]
head(data_merged)
```
# Histogram of the total number of steps taken each day after missing values are imputed

```{r}
steps_grouped_imputed <- data_merged %>% group_by(date) %>% summarise_at(vars(c("steps.x")), list(steps = sum))
hist(steps_grouped_imputed$steps, main="daily steps", breaks=10)
```

# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```{r}
library(lubridate)
steps_avg$weekday <- wday(as.Date(steps_avg$date))
steps_avg$weekend <- steps_avg$weekday==1 | steps_avg$weekday==7
box_plot <- ggplot(steps_avg, aes(x = weekend, y = steps))
box_plot +
    geom_boxplot()
```


